### 2023-03-09 article3.qmd
update

basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues',
        grid.col = 'black') +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  scale_color_brewer(palette = 'Dark2') + theme_bw() +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) #+
  # basemap(limits = c(-55,-40,55,65), bathymetry = F, lat.interval = 2.5,
  #       grid.col = 'black')

@@

The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS \[@Lacour2015-da\]. This is not surprising given that this pair was deployed one degree southward. The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June \[@Lacour2015-da\]. The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD and that it started again in mid-June when the water column was well stratified. However, the Chla satellite concentration for that pair was three times lower than for the first one. Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.

>>>>

p <- basemap(limits = c(-55,-40,55,65), bathymetry = T, lat.interval = 2.5, bathy.style = 'poly_blues')

# Make the graticules:
lims <- attributes(p)$limits 
graticule <- sf::st_graticule(
  c(lims[1], lims[3], lims[2], lims[4]), 
  crs = attributes(p)$proj,
  lon = attributes(p)$map.grid$lon.breaks, 
  lat = attributes(p)$map.grid$lat.breaks
)

p <- p + 
  geom_sf(data = graticule, color = "grey", size = LS(1)) + # graticules
  coord_sf(xlim = lims[1:2], ylim = lims[3:4], # redefine limits
           crs = attributes(p)$proj)

p +
  geom_path(data = argo_map_geo, aes(lon, lat, colour = WMO), linewidth = 1) + 
  geom_point(data = argo_map_geo, aes(x= lon, y = lat, colour = WMO), size = 2) +
  scale_color_brewer(palette = 'Dark2') + theme_bw() +
  geom_point(data = transform_coord(argo_map_first), aes(lon, lat), shape = 17, size = 3) +
  geom_point(data = argo_map_last_geo, aes(lon, lat, colour = WMO), size = 4) +
  geom_point(data = argo_map_last_geo, aes(lon, lat), fill = 'black', size = 4, shape = 1) +
  #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
  guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
                                                  shape = c(NA, NA, NA, NA)))) +
  annotation_scale(location = "br") + 
  annotation_north_arrow(location = "tr", which_north = "true") 
```

@@

The Chla satellite pattern for the pair 6904240-4903634 typically describes the phenology of the spring bloom in the northern part of the LS [@Lacour2015-da]. This is not surprising given that this pair was deployed one degree southward. The other pair was deployed in the bioregion where the phytoplankton biomass slowly develops during May to reach its maximum in June [@Lacour2015-da]. The Chla pattern observed in @fig-chla-mld-satellite could suggest that the development of the phytoplankton biomass was stopped by an increase in the MLD and that it started again in mid-June when the water column was well stratified. However, the Chla satellite concentration for that pair was three times lower than for the first one. Hence, it is possible that the bloom intensity was not sufficient to produce clear particle export signals, at least for BGC-Argo floats in their 10-day vertical sampling configuration.

### 2023-03-09 uvp6_part.qmd
update

The discussion around automatic plankton classification is a key topic in oceanography and the development of convolutional neural networks (CNNs) within that frame has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya]. Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge. Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g. area, eccentricity, grey levels) have outperformed the latter for plankton recognition [@Ellen2019-nh]. Though, it comes at a cost. CNNs are energy consuming hence their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].

@@

The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx]. It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5 (@fig-uvp-pipeline). There are 2 versions: the LP (low power) and the HF (high frequency). The former uses less energy (lower image acquisition rate) and will be deployed on AUVs while the latter is designed to be used on CTD rosettes and cruising AUVs [@Picheral2021-sx].

UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative. For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement. Ultimately, we will train the ER model for the UVP6 by applying observations made on the preliminary UVP5 model.

@@

```{r, classification report XGBoost UVP5}
#| label: tbl-classification-report-uvp5-xgboost-cnn
#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). avg average. avg stands for average.

@@

To build the ER model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e. learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$)). The second step consisted in the search of the optimal number of boosting rounds (i.e. $k$) before the model overfits. For this, we searched for the minimum value of the multi class log loss(also known as logistic loss or cross-entropy loss) evaluated at each boosting round. Using that optimal $k$ value, we then trained the ER model with the entire training set. Therefore, finding the best model consisted in finding the couple ($\eta$, $D_{max}$) that respected the energy constraints and outputed the best global metrics. The latters were determined using a customary grid search (@fig-grid-search) while $\omega$ was chosen empirically (see hereafter).

@@

It is well known that the task of labeling images in supervised machine learning tasks is time consuming and most probably painful when you have millions of images to label. It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree. All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharea*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).

Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (i.e. grid search results shown in @fig-grid-search). In our case, the optimal model hyperparameters were 0.2 for $\eta$ and 5 for D$_{max}$ (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (i.e. loading of the ER model in memory) below 700 ms for energetic reasons. In that configuration, the UVP5 model had an accuracy of 81%, a macro (i.e. label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%. CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.

@@

Quantification is an interesting concept that focuses on the sample instead of the individual. It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon $j$. If the classifier is perfect, $p_{i,j} = 1$ where $i = j$, the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu). However, a perfect (especially in plankton ecology) classifier does not exist. Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biased. This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that P needs to be adapted to the sampling conditions. This reevaluation of $\pmb{P}$ is due to what is defined as a data set shift where part of the unseen data follows a different distribution than the one from the training data [@Orenstein2020-jl; @Moreno-Torres2012-oq]). It is especially true for ER systems deployed on AUVs where planktonic communities change throughout the year. Thus, we tested different quantification methods (CC, adjusted count (AC), threshold count (TC), probabilistic adjusted count (PAC) and scaled probability adjusted count (SPAC), see @Gonzalez2017-tr for a review on those quantification methods) assuming that the estimation of $\pmb{P}$ was accurate enough (\> 3M images in our training set) and that the global coverage of UVP5 profiles in the training set was sufficient to cover various planktonic distributions from the test set.

@@

Increasing the number of detritus classes was also considered in the hope of clustering them in the features space. Detritus display a variety of size, shape and grey levels so it must be possible to cluster them and therefore improve the classification score. For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier. @fig-grid-search clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision. In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus. Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.

<!-- ![Confusion matrix (normalized on true labels) of the a UVP5 XGBoost model with 5 classes of detritus.](image/uvp6/Muvpec_367314a7_confusion_matrix.jpeg){#fig-confusion-matrix-uvp6-ndetritus} -->
```{r, confusion matrix N detritus}
#| label: fig-confusion-matrix-uvp6-ndetritus
#| fig-cap: Confusion matrix (normalized on true labels) of a UVP5 XGBoost model with 5 classes of detritus.

@@

```
#### Application to the UVP6 data set

@@

The specifications of the final UVP6 ER model are the following: $\eta$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms. Note that the CV initially found an optimal number of boosting rounds of 555 for a loading time of 101 ms). However, we observed that the log loss curve had already reached a plateau hence we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, decreasing a bit more the energy consumption of the UVP6 on board AUVs.

@@

```{r, uvp6 BGC application}
#| label: fig-bgc-application-uvp6
#| fig-cap : Concentrations of 4 classes (3 biological and detritus) estimated with data acquired from the UVP6 on board the BGC-Argo float (WMO 4903634) on the 15$^{th}$ of June 2022 in the Labrador Sea at latitude 59.27° N and longitude -50.56° E. Red points represent positive concentrations. Dotted lines represent the depth of the deep chlorophyll maximum.

@@

The objective of this paper was to develop a zooplankton classification model under strict energy constraints for the UVP6 to be deployed on BGC-Argo floats. It was clear at the beginning that the use of powerful energy consuming CNNs had to be discarded and that we needed to build a satisfactory features-based model instead. To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights). In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy. Afterwards, we applied the data processing pipeline and the recommendations from the UVP5 data set analysis on the UVP6 data set to create the final ER model. This model contains 20 labels, 16 biological and 4 non-living classes, with a total loading time of 73 ms. The initial upper limit for the loading time was 700 ms therefore we are saving precious energy to extend the lifetime of BGC-Argo floats. Finally, we showed the CM and CR of the final model and how to use it to interpret some UVP6 taxonomic profiles observed by the float WMO 4903634 during the spring bloom in the southern part of the Labrador Sea.

>>>>

The discussion around automatic plankton classification is a key topic in oceanography and the development of convolutional neural networks (CNNs) within that frame has already lead to promising results [@Luo2018-oo; @Cui2018-nt; @Dai2017-ex; @Campbell2020-cq; @Schroder2019-um; @Bochinski2019-af; @Lee2016-jl; @Cheng2019-ar; @Guo2021-xi; @Ellen2019-nh; @Orenstein2017-ya]. Nevertheless, real-time image processing done at sea is far less mentioned to our knowledge. Usually, images are brought back to the lab and classified with CNNs which, in comparison with features-based classifiers, that is machine learning algorithms using a set of features extracted from the image (e.g. area, eccentricity, gray levels) have outperformed the latter for plankton recognition [@Ellen2019-nh]. Though, it comes at a cost. CNNs are energy consuming hence their implementation for embedded applications can be limited by power consumption [@Biswas2018-uw; @Khabbazan2019-oi; @Oh2017-kj].

@@

The UVP6 is the miniaturized version of the UVP5 [@Picheral2021-sx]. It is smaller, it uses less energy and it has been specifically developed to be mounted on AUVs such as BGC-Argo floats with the aim of providing data of similar quality than from the UVP5. There are 2 versions: the LP (low power) and the HF (high frequency). The former uses less energy (lower image acquisition rate) and will be deployed on AUVs while the latter is designed to be used on CTD rosettes and cruising AUVs [@Picheral2021-sx].

UVP5 Standard Definition (SD) profiles have been collected since 2008 at a global scale (@fig-map-of-profiles) whereas the UVP6 has been deployed very recently in specific locations and thus lacks spatial coverage to be fully representative. For this reason, UVP5 data will be used to create a preliminary ER model and explore ways of improvement. Ultimately, we will train the ER model for the UVP6 by applying the lessons learnt from the preliminary UVP5 model.

@@

```{r, classification report XGBoost UVP5}
#| label: tbl-classification-report-uvp5-xgboost-cnn
#| tbl-cap: Precision, recall and f1-score for each class and globally (lightgray background) in the UVP5 model (XGBoost and CNN). The abbreviation avg stands for average.

@@

To build the ER model, the first step was a threefold cross-validation using XGBoost parameterized with a set of given hyperparameters (i.e. learning rate ($\eta$), maximum depth of a tree ($D_{max}$) and weight sensitivity ($\omega$)). The second step consisted in the search of the optimal number of boosting rounds (i.e. $k$) before the model overfits. For this, we searched for the minimum value of the multi class log loss (also known as logistic loss or cross-entropy loss) evaluated at each boosting round. Using that optimal $k$ value, we then trained the ER model with the entire training set. Therefore, finding the best model consisted in finding the couple ($\eta$, $D_{max}$) that respected the energy constraints and outputed the best global metrics. The latters were determined using a customary grid search (@fig-grid-search) while $\omega$ was chosen empirically (see hereafter).

@@

It is well known that the task of labeling images in supervised machine learning tasks is time consuming and most probably painful when you have millions of images to label. It is also very frustrating to see that one class (largely) dominates when you are interested in the small ones and that some of them are so scarce that you need to merge them with their siblings (at the very best) on the phylogenetic tree. All in all, we found - after multiple trials and errors - 11 classes fulfilling a first compromise between classification performances and ecological importance: 7 biological classes (*Acantharia*, *Aulacanthidae*, *Aulosphaeridae*, *Copepoda*, *Trichodesmium*, darksphere and other\<living) and 4 non-living classes (detritus, bubble, fiber and artefact).

Once the number of classes is known, hyperparameters can be tuned to find the optimal couple for the classification results (i.e. grid search results shown in @fig-grid-search). In our case, the optimal model hyperparameters were 0.2 for $\eta$ and 5 for D$_{max}$ (also with only 1 class of detritus, see Discussion) to have both a good weighted living (or biological) precision and recall and a loading time (i.e. loading of the ER model in memory) below 700 ms for energetic reasons. In that configuration, the UVP5 model had an accuracy of 81%, a macro (i.e. label imbalance not taken into account) precision and recall of, respectively, 64% and 69% and a weighted living precision and recall of, respectively, 60% and 65%. The CV found an optimal number of 850 boosting rounds before over-fitting and the expected loading time is 382 ms.

@@

Quantification is a concept that focuses on the sample instead of the individual. It was already shown in @Solow2001-jc, using plankton images from the Video Plankton Recorder, that the estimation of taxonomic abundances could be improved by computing the confusion matrix $\pmb{P}$ where $p_{i,j}$ is the probability that a taxon $i$ is classified as taxon $j$. If the classifier is perfect, $p_{i,j} = 1$ where $i = j$, the quantification resumes to a simple Classify and Count method (CC, see @Forman2005-eu). However, a perfect (especially in plankton ecology) classifier does not exist. Therefore, quantification is a tool to better estimate the abundance of planktonic groups while perfectly knowing that the classifier is biased. This quantification method was successively applied in the paper of @Sosik2007-cj on a 2 month data set although it is very likely that P needs to be adapted to the sampling conditions. This reevaluation of $\pmb{P}$ is due to what is defined as a data set shift where part of the unseen data follows a different distribution than the one from the training data [@Orenstein2020-jl; @Moreno-Torres2012-oq]). It is especially true for ER systems deployed on AUVs where planktonic communities change throughout the year. Thus, we tested different quantification methods (CC, adjusted count (AC), threshold count (TC), probabilistic adjusted count (PAC) and scaled probability adjusted count (SPAC), see @Gonzalez2017-tr for a review on those quantification methods) assuming that the estimation of $\pmb{P}$ was accurate enough (\> 3M images in our training set) and that the global coverage of UVP5 profiles in the training set was sufficient to cover various planktonic distributions from the test set.

@@

Increasing the number of detritus classes was also considered in the hope of clustering them in the features space. Detritus display a variety of size, shape and gray levels so it must be possible to cluster them and therefore improve the classification score. For this, we applied a classic K-Means algorithm [@Hartigan1979-oz] on detritus cleaned from likely outliers using an Isolation Forest Algorithm [@Liu2012-xz] and we executed a grid search as described earlier. @fig-grid-search clearly shows that using more than 1 class of detritus drastically decreases the global accuracy of the model, decreases the weighted living recall and only increases by 1% the weighted living precision. In addition, the confusion matrix (@fig-confusion-matrix-uvp6-ndetritus) shows that it does not improve the ER model because there is a high confusion between subgroups of detritus. Essentially, it seems that detritus form a continuum in the features space and that more research should be done on the choice of features, clustering algorithms and related methods to better differentiate them in separate entities, which is not the purpose of this study.

<!-- ![Confusion matrix (normalized on true labels) of the a UVP5 XGBoost model with 5 classes of detritus.](image/uvp6/Muvpec_367314a7_confusion_matrix.jpeg){#fig-confusion-matrix-uvp6-ndetritus} -->

```{r, confusion matrix N detritus}
#| label: fig-confusion-matrix-uvp6-ndetritus
#| fig-cap: Confusion matrix (normalized on true labels) of a UVP5 XGBoost model with 5 classes of detritus.

@@

#### Application to the UVP6 data set

@@

The specifications of the final UVP6 ER model are the following: $\eta$: 0.2, $D_{max}$: 3, $\omega$: 0.25, number of boosting rounds: 400 for a total loading time of 73 ms. Note that the CV initially found an optimal number of boosting rounds of 555 for a loading time of 101 ms). However, we observed that the log loss curve had already reached a plateau hence we were able to decrease the size of the model without losing much (\< 1%) in biological precision or recall, decreasing a bit more the energy consumption of the UVP6.

@@

```{r, uvp6 BGC application, fig.width=6, fig.height=5}
#| label: fig-bgc-application-uvp6
#| fig-cap : Concentrations of 4 classes (3 biological and detritus) estimated with data acquired from the UVP6 on board the BGC-Argo float (WMO 4903634) on the 15$^{th}$ of June 2022 in the Labrador Sea at latitude 59.27° N and longitude -50.56° E. Red points represent positive concentrations. Dotted lines represent the depth of the deep chlorophyll maximum.

@@

The objective of this paper was to develop a zooplankton classification model under strict energy constraints for the UVP6 to be deployed on BGC-Argo floats. It was clear at the beginning that the use of powerful energy consuming CNNs had to be disregarded and that we needed to build a satisfactory features-based model instead. To this aim, we firstly used XGBoost on UVP5 data, a more spatially and temporally representative data set than the recent UVP6 data set to create the data processing pipeline (classes aggregation, hyperparameter tuning, model training, see also @fig-uvp-pipeline) and investigate ways to improve classification scores (quantification, detritus subgroups, label weights). In order to see how the model could be improved, we compared the XGBoost model with results from a CNN not limited by energy. Afterwards, we applied the data processing pipeline and the recommendations from the UVP5 data set analysis on the UVP6 data set to create the final ER model. This model contains 20 labels, 16 biological and 4 non-living classes, with a total loading time of 73 ms. The initial upper limit for the loading time was 700 ms therefore we are saving precious energy to extend the lifetime of BGC-Argo floats. Finally, we showed the CM and CR of the final model and how to use it to interpret some UVP6 taxonomic profiles observed by the float WMO 4903634 during the spring bloom in the southern part of the Labrador Sea.
